Economic Drivers, Immigration Dynamics, and Housing Starts: Their Collective Influence on Housing Price Index

### 2024-04-18 paper/paper.qmd
added model files git comm

thanks: "Code and data are available at: https://github.com/hoodanav/Housing_Market_Analysis/tree/main"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(palmerpenguins)
library(here)
library(ggplot2)
library(scales)

@@

With the complexity of the housing market influenced by numerous factors, keeping track of its evolution proves challenging. This analysis aims to understand the dynamics of the housing market, exploring the impact of economic variables and identifying significant contributors. By examining economic indicator changes, we seek to uncover their significance and understand correlations among variables and their market impact.

@@

### Immigration Numbers
The yearly immigration numbers were obtained from Statista [@immigrants], and was originally sourced from Statistics Canada. This data reports the number of immigrants Canada welcomes in each year from 2000-2023. The data was collected and sourced from a survey done by Statistics Canada. A visual of these numbers is seen in @fig-immigration-plot.

@@

## Data Cleaning 
Simple data cleaning operations were applied to achieve a cohesive merged data set for analysis. We used R [@citeR] for data cleaning and processing, using packages like tidyverse [@tidy] for data manipulation and janitor [@jan] for cleaning column names. Other packages used includes `ggplot2` [@ggplot], `dplyr` [@dp], `readr` [@read], `tibble` [@tib], `janitor` [@jan], `knitr` [@knit], `ggbeeswarm` [@gg], `ggrepel` [@repel], `kableExtra`[@kable], `readxl`[@readxl], `rstanarm`[@rstan], `modelsummary`[@model] and `here` [@here]. All these different data sets were cleaned in a different way to match the final representation of the cleaned data.

@@

## Data Modifications
Overall, the values were often reported at different times, for consistency we chose to use a yearly metric. We averaged upon quarterly or monthly reported data to gain an averaged yearly value.

@@

```{r}
#| label: hpivs-plot
#| fig-cap: Number of immigrants in Canada from 2000 to 2023
#| echo: false
#| warning: false
#| message: false
# Scatterplot of HPI against each economic indicator
library(ggplot2)

@@

House_Price_Index = β_0 + β_1 × Inflation_Rate + β_2 × Interest_Rate + β_3 × Unemployment_Rate + β_4 × Immigration_numbers + β_5 × Housing_Starts + ε

@@

## Results {#sec-data3}

@@

@@

Since Our model could not grasp the nature of its data and showed a very minute relationship. To show some key insights, transforming these variables to a concise unit of change could help highlight their importance, or additionally using already aggregated data could help the regression model uncover associations more meaningfully.

>>>>

thanks: "Code and data are available at: https://github.com/hoodanav/Housing_Market_Analysis/tree/main"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

@@

library(tidyverse)
library(palmerpenguins)
library(here)
library(ggplot2)
library(scales)
library(dplyr)
library(knitr)
library(kableExtra)

@@

With the complexity of the housing market influenced by numerous factors, keeping track of its evolution proves challenging. The estimand of this paper is to find the associations the of economic variables, the supply demand variables represented by housing starts and immigration numbers to the house price index. By examining economic indicator changes, we seek to uncover their significance and understand correlations among variables and their market impact.

@@

### Immigration Numbers
The yearly immigration numbers were obtained from Statista [@immigrants], and was originally sourced from Statistics Canada. This data reports the number of immigrants Canada welcomes in each year from 2000-2023 in thousands. The data was collected and sourced from a survey done by Statistics Canada. A visual of these numbers is seen in @fig-immigration-plot.

@@

```{r}
# Load necessary libraries
library(dplyr)
library(knitr)
library(kableExtra)

# Display the first 6 rows of the merged data as a table using kable with proper styling
kable(head(merged_data, 6), format = "latex", caption = "Merged Data Table") %>%
  kable_styling(latex_options = "scale_down") %>%
  as.character() %>%
  cat("\n\\centering")  # Add LaTeX centering command

```
## Data Cleaning 
Simple data cleaning operations were applied to achieve a cohesive merged data set for analysis. We used R [@citeR] for data cleaning and processing, using packages like tidyverse [@tidy] for data manipulation and janitor [@jan] for cleaning column names. Other packages used includes `ggplot2` [@ggplot], `dplyr` [@dp], `readr` [@read], `tibble` [@tib], `janitor` [@jan], `knitr` [@knit], `ggrepel` [@repel], `kableExtra`[@kable], `readxl`[@readxl], `rstanarm`[@rstan], `modelsummary`[@model] and `here` [@here]. All these different data sets were cleaned in a different way to match the final representation of the cleaned data.

@@

## Data Modifications
Overall, the values were often reported at different times, for consistency we chose to use a yearly metric. We averaged upon quarterly or monthly reported data to gain an averaged yearly value. We then created a merged datadframe joined by the year variable and achieved our cleaned data as seen in @tbl-clean-data.

@@

```{r}
#| label: fig-hpivs-plot
#| fig-cap: Number of immigrants in Canada from 2000 to 2023
#| echo: false
#| warning: false
#| message: false
# Scatterplot of HPI against each economic indicator
library(ggplot2)

@@

\[ \text{House\_Price\_Index} = \beta_0 + \beta_1 \times \text{Inflation\_Rate} + \beta_2 \times \text{Interest\_Rate} + \beta_3 \times \text{Unemployment\_Rate} + \beta_4 \times \text{Immigration\_numbers} + \beta_5 \times \text{Housing\_Starts} + \epsilon \]

@@

# Results {#sec-data3}

@@

## Interest Rates and HPI
The model suggests interest rate as a significant predictor of HPI

@@

Since our model could not grasp the nature of its data and showed a very minute relationship. To show some key insights, transforming these variables to a concise unit of change could help highlight their importance, or additionally using already aggregated data could help the regression model uncover associations more meaningfully.
